---
title: "Ad Clicking"
author: "Jim Reilly"
date: "February 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r clear the global env}
rm(list=ls())
```

```{r message=FALSE}
library(dplyr)
library(ggplot2)
library(rpart)
library(rpart.plot)
library(caret)
library(randomForest)
```

###1) Loading the data

Our data is a .csv with 1,000,000 entries. We must also convert a couple columns to factors.

```{r load the data}
ads <- read.csv(file = 'data/criteo_ad_click_sampled.csv', header = FALSE)
```

###2) Transform the data

We need to transform some of our columns before we can build regressions. I will convert the label (column V1) to a factor so that we can build a binary classifier later.

```{r mutate the dataset}
ads <- ads %>% mutate(V1 = factor(V1))

head(ads)
```

###3) Exploratory Analysis

We will look at some preliminary features to decide what variables might be good predictors and to better understand the available data

```{r summary of the set}
clicks <- summary(ads$V1)
```
We have `r clicks[1]` no clicks and `r clicks[2]` clicks. Since we care more about the clicks than the no clicks we might consider some oversampling as described in the text.

We also have a very large quantity of data. I will produce my visualizations off of a sample for performance concerns.

```{r random sample with seed}
set.seed(4747) # set seed for reproducability

ads.sample.index <- sample(1:nrow(ads), 10000)

ads.sample <- ads[ads.sample.index, ]
```

```{r visualization of V2}
v2.plot <- ggplot(ads.sample, aes(x = V1, y = V2))

v2.plot + geom_boxplot()
v2.plot + geom_jitter()
```

```{r visualization of V3}
v3.plot <- ggplot(ads.sample, aes(x = V1, y = V3))

v3.plot + geom_boxplot()
v3.plot + geom_jitter()
```

```{r visualization of V4}
v4.plot <- ggplot(ads.sample, aes(x = V1, y = V4))

v4.plot + geom_boxplot()
v4.plot + geom_jitter()
```

```{r visualization of V5}
v5.plot <- ggplot(ads.sample, aes(x = V1, y = V5))

v5.plot + geom_boxplot()
v5.plot + geom_jitter()
```


```{r visualization of V6}
v6.plot <- ggplot(ads.sample, aes(x = V1, y = V6))

v6.plot + geom_boxplot()
v6.plot + geom_jitter()
```

```{r visualization of V7}
v7.plot <- ggplot(ads.sample, aes(x = V1, y = V7))

v7.plot + geom_boxplot()
v7.plot + geom_jitter()
```

```{r visualization of V8}
v8.plot <- ggplot(ads.sample, aes(x = V1, y = V8))

v8.plot + geom_boxplot()
v8.plot + geom_jitter()
```

```{r visualization of V9}
v9.plot <- ggplot(ads.sample, aes(x = V1, y = V9))

v9.plot + geom_boxplot()
v9.plot + geom_jitter()
```

```{r visualization of V10}
v10.plot <- ggplot(ads.sample, aes(x = V1, y = V10))

v10.plot + geom_boxplot()
v10.plot + geom_jitter()
```

```{r visualization of V11}
v11.plot <- ggplot(ads.sample, aes(x = V1, y = V11))

v11.plot + geom_boxplot()
v11.plot + geom_jitter()
```

```{r visualization of V12}
v12.plot <- ggplot(ads.sample, aes(x = V1, y = V12))

v12.plot + geom_boxplot()
v12.plot + geom_jitter()
```

```{r visualization of V13}
v13.plot <- ggplot(ads.sample, aes(x = V1, y = V13))

v13.plot + geom_boxplot()
v13.plot + geom_jitter()
```

```{r visualization of V14}
v14.plot <- ggplot(ads.sample, aes(x = V1, y = V14))

v14.plot + geom_boxplot()
v14.plot + geom_jitter()
```

```{r visualization of V15}
v15.plot <- ggplot(ads.sample, aes(x = V1, y = V15))

v15.plot + geom_boxplot()
v15.plot + geom_jitter()
```

```{r visualization of V16}
v16.plot <- ggplot(ads.sample, aes(x = V1, y = V16))

v16.plot + geom_boxplot()
v16.plot + geom_jitter()
```

```{r visualization of V17}
v17.plot <- ggplot(ads.sample, aes(x = V1, y = V17))

v17.plot + geom_boxplot()
v17.plot + geom_jitter()
```

```{r visualization of V18}
v18.plot <- ggplot(ads.sample, aes(x = V1, y = V18))

v18.plot + geom_boxplot()
v18.plot + geom_jitter()
```

```{r visualization of V19}
v19.plot <- ggplot(ads.sample, aes(x = V1, y = V19))

v19.plot + geom_boxplot()
v19.plot + geom_jitter()
```

```{r visualization of V20}
v20.plot <- ggplot(ads.sample, aes(x = V1, y = V20))

v20.plot + geom_boxplot()
v20.plot + geom_jitter()
```

```{r visualization of V21}
v21.plot <- ggplot(ads.sample, aes(x = V1, y = V21))

v21.plot + geom_boxplot()
v21.plot + geom_jitter()
```

```{r visualization of V22}
v22.plot <- ggplot(ads.sample, aes(x = V1, y = V22))

v22.plot + geom_boxplot()
v22.plot + geom_jitter()
```

V5-V10 and V14 show a pattern where there seems to be a higher median value for the `0` target vs. the `1`. I would assume that these will be identified as better predictors than some of the other choices. Of the factors, V15 - V18 seem to have a skew in the number of points belonging to each class for some factor values. The remaining factors have too low variance between the two targets or if there is a major difference the sample size in that factor appears small on the jitter plot (which is significant for 100,000 records). In order to best reason with the factors individually I convert them to dummy variables with a 0 or 1 encoding.

```{r ads logistic}
ads.sample <- fastDummies::dummy_cols(ads.sample)


ads.sample.glm <- glm(V1 ~ ., data = ads.sample, family = "binomial")
ads.sample.glm <- step(ads.sample.glm,  direction = "backward")
```

```{r test}
default.ct <- rpart(V1 ~ ., data = ads.sample, method = "class")
prp(default.ct, type = 1, extra = 1, under = TRUE, split.font = 1, varlen = -10)
default.ct.pred <- predict(default.ct, ads.sample, type = "class")
confusionMatrix(default.ct.pred, ads.sample$V1)

default.ct <- rpart(V1 ~ ., data = ads.sample, method = "class")
default.ct.pred <- predict(default.ct, ads, type = "class")
confusionMatrix(default.ct.pred, ads$V1)

ads.ct <- rpart(V1 ~ ., data = ads, method = "class")
ads.ct.pred <- predict(ads.ct, ads, type = "class")
confusionMatrix(ads.ct.pred, ads$V1)
prp(ads.ct, type = 1, extra = 1, under = TRUE, split.font = 1, varlen = -10)

ads.sample.rf <- randomForest(V1 ~ ., data = ads.sample, ntree = 100, mtry = 4, nodesize = 5, importance = TRUE)
```
