---
title: "Milwaukee Weather"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

This notebook will walk through the exercise of opening a dataset of daily temperature and percipitation to explore introductory visualization models.


###1) What is the form of the data?

```{r load the data}
#Load the data
data = read.csv("data/milwaukee_weather.csv")

nrows <- nrow(data)
ncolumns <- ncol(data)
head(data)
```

Our dataset has `r nrows` records of weather and `r ncolumns` variables on each record. We will use these records to make various plots below

###2) Scatter plot

Lets view the average temperature over time. I am choosing a scatter plot for each record so that I can best view patterns over time of the year. I will be using ggplot2 as the plotting library to make these visuals.

```{r install ggplot, include=FALSE}
#Install ggplot2
install.packages("ggplot2")

```

```{r, average temp plot}
library(ggplot2)

ggplot(data, aes(date, avg_temp)) + geom_point(size=1)


```