---
title: "Election Data"
author: "Jim Reilly"
date: "February 4, 2019"
output: html_document
---

Load the required libraries, `ggplot2`, `tidyverse`, and `readxl`

```{r libraries, message = FALSE}
library(ggplot2)
library(tidyverse)
library(readxl)
library(cellranger)
library(dplyr)
library(purrr)
library(maps)
library(sf)
library(sets)
```


## Datasets

The datasets for this exercise are two Excel spreadsheets (`.xlsx`) across multiple tables with federal election data from the 2016 Senate, House, and Presidential election. As well as election results for the previous 5 presedential elections.

Using `readxl`, load the dataset into memory

```{r loading the data}
pop_est_2016 <- read_excel("data/nst-est2016-01.xlsx", range="J10:J60", col_names = FALSE)
states <- read_excel("data/nst-est2016-01.xlsx", range="A10:A60", col_names = FALSE)

# Remove a '.' character infront of each string from the read file
states <- states %>%
  purrr::map(function(x) substring(x, first = 2))
```

The data from an excel spreadsheet. Looking at the file, only 2 ranges were required to plot the estimated population per state. To make graphing easier, its best to arrange this data into a single dataframe before passing it to `ggplot`

```{r preprocessing the data}
# Construct a dataframe from the 2 lists and label the columns
pop_per_state <- map2_dfr(states, pop_est_2016, ~ tibble(A = .x, B = .y))
colnames(pop_per_state) <- c("State", "Population")

#reorder the factor levels by population, descending for graphing
pop_per_state[order(-pop_per_state$Population),]
pop_per_state$State <- factor(pop_per_state$State, levels = pop_per_state$State[order(-pop_per_state$Population)])

#a variable from 1.0 to 0 representing the ratio between a states size and the largest state for coloration in a map plot
pop_per_state$Index <- pop_per_state$Population %>%
  purrr::map_dbl(function(pop) pop / max(pop_per_state$Population))

# plot a col chart of the population estimate per state in descending order
ggplot(data = select(pop_per_state, State, Population), aes(x = State, y = Population)) + geom_col() + theme(axis.text.x = element_text(angle = 90, hjust = 1))

pop_map <- maps::map("state", plot = FALSE, fill = TRUE)
pop_map <- sf::st_as_sf(pop_map)
pop_map <- pop_map %>% filter(ID != 'district of columbia')
pop_map[c('ID')] <- lapply(pop_map[c('ID')], as.character)

# combine with population data
pop_map <- pop_map %>% 
  left_join(pop_per_state %>% mutate(State = tolower(State)),by = c("ID" = "State")) 

ggplot(pop_map) + geom_sf(aes(fill = Index)) + scale_fill_viridis_c(option = "plasma")

```

California has the highest estimated population and Wyoming has the smallest. Interestingly, if the District of Columbia became a state it would be more populous that 2 existing states (Wyoming and Vermont)

Using the gradient map graph, its easy to see how relative to california there is virtually no discernable difference between most of the states' population. It would take a transform such as `sqrt` or `log` to visualize more meaningful population discrepency


```{r load percentage of votes for candidate}
results_2016 <- read_excel("data/federalelections2016.xlsx", sheet = "2016 Pres General Results") %>%
  select(state = STATE, last_name = 'LAST NAME', party = PARTY, votes = 'GENERAL RESULTS', vote_percentage = 'GENERAL %') %>%
  filter(last_name != 'Scattered') %>%
  filter(party != 'W') %>%
  dplyr::mutate_if(is.character, as.factor)

results_2016$votes_per_candidate <- apply(results_2016[,c("last_name", "votes")], 1, function(x) tuple(x[1], x[2]))

  
split(results_2016$votes, results_2016$state)
```

```{r state plot}
# wrangle presidential election data
library(dplyr)
library(politicaldata)

pres <- politicaldata::pres_results

pres <- pres %>% 
  filter(year%in% 2008:2016)


pres$state_name = tolower(state.name[match(pres$state,state.abb)]) # we need the full name of the state to join with the map


# get map data
library(maps)
library(sf)
pres_map <- maps::map("state", plot = FALSE, fill = TRUE) # acquirig the data frame

pres_map <- sf::st_as_sf(pres_map) # changing the lat-long format to a simple feature object

pres_map <- pres_map %>% filter(ID != 'district of columbia')

pres_map[c('ID')] <- lapply(pres_map[c('ID')], as.character)

# combine with election data
pres_map <- pres_map %>% 
  left_join(pres,by = c("ID" = "state_name")) # joining the map with the political data

# plotting with ggplot2
library(ggplot2)
gg <- ggplot(pres_map,
             aes(fill=dem-rep>0),
             col='black') +
  geom_sf(aes(alpha=abs(dem-rep))) +
  coord_sf(crs = "+proj=aea +lat_1=25 +lat_2=50 +lon_0=-100",ndiscr = 0) + # change the projection from mercator (blegh)
  scale_fill_manual(values=c("TRUE"="blue","FALSE"="red")) +
  scale_alpha(range=c(0.1,1)) +
  facet_wrap(~year) +
  theme_void() +
  theme(legend.position='none')

gg

#ggsave(plot = gg,filename = 'us_election_2008_2012_2016.png',width=12,height=3,units='in')

```