---
title: "Cereal Clustering"
author: "Jim Reilly"
date: "April 1, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include = FALSE, message = FALSE}
library(caret)
library(dplyr)
```

## Data

I load the cereals while also removing rows that have missing values

```{r load the data}
cereal <- na.omit(read.csv("data/Cereals.csv"))
head(cereal)
```

## Normalize the data

Normalize the data with caret `preProcess` function
```{r normalize}
#find what columns are numeric
numerics <- grep(paste(c("numeric", "integer"), collapse="|"), lapply(cereal,class))

#define the normalization function
normalize <- function(x) { (x - min(x)) / (max(x) - min(x)) }

#apply the function to the numeric columns
cereal.normal<- apply(cereal[,numerics], MARGIN = 2, FUN = normalize) 
row.names(cereal.normal) <- cereal$name
```

## Compute distances
```{r distances}
cereal.dist.euc <- dist(cereal.normal, method = "euclidean")
```

## Clustering

Single hierarchical and Complete hierarchical clustering

```{r clustering}
cereal.complete <- hclust(cereal.dist.euc, method = "complete")
plot(cereal.complete, hang = -1, ann = FALSE)

cereal.single <- hclust(cereal.dist.euc, method = "single")
plot(cereal.single, hang =  -1, ann = FALSE)
```
## Make clusters
The complete hierarchical cluster looks much more evenly distributed.

Based on the complete hierarchical clustering, I think 7 clusters is a good amount based on the dendrogram

```{r cutree}
cereal.clusters = cutree(cereal.complete, k = 7)

cereal.clusters.df <- data.frame(name = names(cereal.clusters), cluster = cereal.clusters) %>% arrange(cluster)
cereal.clusters.df
```

## Centroid analysis

```{r centroid means}
#find the medoid in a cluster 'i' (from https://www.biostars.org/p/13143/)
centroid.means = function(i, data, clusters) {
  numerics <- grep(paste(c("numeric", "integer"), collapse="|"), lapply(cereal,class))
  j <- (clusters == i)
  colMeans(data[j,numerics])
}

sapply(unique(cereal.clusters), centroid.means, cereal, cereal.clusters)
```
